---
title: "vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(Lab6)
```


```{r}

suppressWarnings(RNGversion(min(as.character(getRversion()),"3.5.3")))

set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 1000000
knapsack_objects <- data.frame(
w=sample(1:4000, size = n, replace = TRUE),
v=runif(n = n, 0, 10000)
)

```



## Brute force search
Question: How much time does it takes to run the algorithm for n = 16 objects?
```{r}

# system.time(brute_force_knapsack(x = knapsack_objects[1:16,], W = 3500))

```


## Dynamic programming
Question: How much time does it takes to run the algorithm for n = 500 objects?
```{r}

# system.time(dynamic_knapsack(x = knapsack_objects[1:500,], W = 3500))

```


## Greedy heuristic
Question How much time does it takes to run the algorithm for n = 1000000 objects?
```{r}

# system.time(greedy_knapsack(x = knapsack_objects, W = 3500))

```


## Profile your code and optimize your code
Question: What performance gain could you get by trying to improving your code?
```{r}

```


## (*) Parallelize brute force search
Question: What performance gain could you get by parallelizing brute force search?
```{r}

# system.time(brute_force_knapsack(x = knapsack_objects[1:12,], W = 2000))
# system.time(brute_force_knapsack(x = knapsack_objects[1:12,], W = 2000, parallel = TRUE))

```


