---
title: "vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}

library(Lab6)
library(profvis)

```


```{r}

# create data set
suppressWarnings(RNGversion(min(as.character(getRversion()),"3.5.3")))

set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 1000000
knapsack_objects <- data.frame(
w=sample(1:4000, size = n, replace = TRUE),
v=runif(n = n, 0, 10000)
)

```



## Brute force search
Question: How much time does it takes to run the algorithm for n = 16 objects?
```{r}

system.time(brute_force_knapsack(x = knapsack_objects[1:16,], W = 3500))

```


## Dynamic programming
Question: How much time does it takes to run the algorithm for n = 500 objects?
```{r}

system.time(dynamic_knapsack(x = knapsack_objects[1:500,], W = 3500))

```


## Greedy heuristic
Question How much time does it takes to run the algorithm for n = 1000000 objects?
```{r}

system.time(greedy_knapsack(x = knapsack_objects, W = 3500))

```


## Profile your code and optimize your code

Optimizing code such that it runs faster is an iterative process:\newline

1. Find biggest bottleneck (slowest part of the code).\newline
2. Try to eliminate it (may not successful but that's okay). \newline
3. Repeat until code is "fast enough". \newline


There exists a number of techniques to improve a code. Some of them are: \newline

1. Look for existing solutions.\newline
2. Do less work.\newline
3. Vectorise.\newline
4. Parallelise.\newline
5. Avoid copies.\newline
6. Byte-code compile (Wickham, 2019). 



### Brute force alogrithm
```{r}

x = knapsack_objects[1:15,]
W = 3500

profvis({
  
  # check input variables
  stopifnot(is.data.frame(x), x>0, length(x)==2, names(x) == c("w","v"))
  stopifnot(is.numeric(W), W>0, length(W)==1)
  
  # if(!missing(parallel) && !is.logical(parallel)){
  #   stop("variable parallel must be logical")
  # }
  
  # initialize starting values
  val <- 0
  maxval <- 0
  n <- nrow(x)
  elements <- length(n)
  
  lapply(1:n, function(i){
      comb <- utils::combn(n, i) # all the combinations of i from 1 to n
      k <- 1
      while(k <= ncol(comb)){ 
        if (W > sum(x$w[comb[,k]])){
          val <- sum(x$v[comb[,k]])
          if (val > maxval){
            elements <<- comb[,k] # save elements of that combination
            maxval <<- val # save the max value found
          }
        }
        k <- k+1
      }
    })
    result <- list(value = round(maxval), elements=elements)
})

```
The lapply and the while-loop take the most time. These points can be optimized through the use of  parallelization, which is implemented in the regualar function (as the "*" exercise).


### Dynamic alogrithm
```{r}

x = knapsack_objects[1:800,]
W = 3500

profvis({
  
  # check input variables
  stopifnot(is.data.frame(x), x>0, length(x)==2, names(x) == c("w","v"))
  stopifnot(is.numeric(W), W>0, length(W)==1)
  
  
  # initialize starting values
  n <- nrow(x)
  val <- x$v
  wei <- x$w

  m <- matrix(nrow = n+1, ncol = W+1)
  m[1,] <- 0
  m[,1] <- 0
  
  for(i in 1:n){    
    for(c in 0:W){ # c = capacity of knapsack
      
      if(wei[i] > c){
        m[i+1,c+1] <- m[i,c+1]
      } else {
        m[i+1,c+1] <- max(m[i,c+1], m[i,c+1 -wei[i]] + val[i])
      }
    }
  } 
  
  c <- c+1  
  i <- which.max(m[,c])
  elements <- length(n)
  k <- 1
  elements[k] <- i-1
  
  while(m[i,c]!=0 && c!=1 && i!=0){
    k <- k+1
    c <- (c-wei[i-1])
    i <- which(m[,c] == m[i-1,c])[1]
    elements[k] <- i-1
  }
  
  value <- round(m[n+1,W+1])
  elements <- sort(elements[which(elements>0)])
  
  result <- list(value=value, elements=elements)
})

```
As expected the for-loops and the while-loop take most of the computation time. Furthermore the sort() function is also time consuming. 

### System.time Dynamic algorithm comparison
```{r}

x = knapsack_objects[1:800,]
W = 3500

# system.time of the dynamic algorithm
system.time({
  
  # check input variables
  stopifnot(is.data.frame(x), x>0, length(x)==2, names(x) == c("w","v"))
  stopifnot(is.numeric(W), W>0, length(W)==1)
  
  
  # initialize starting values
  n <- nrow(x)
  val <- x$v
  wei <- x$w

  m <- matrix(nrow = n+1, ncol = W+1)
  m[1,] <- 0
  m[,1] <- 0
  
  for(i in 1:n){    
    for(c in 0:W){ # c = capacity of knapsack
      
      if(wei[i] > c){
        m[i+1,c+1] <- m[i,c+1]
      } else {
        m[i+1,c+1] <- max(m[i,c+1], m[i,c+1 -wei[i]] + val[i])
      }
    }
  } 
  
  c <- c+1  
  i <- which.max(m[,c])
  elements <- length(n)
  k <- 1
  elements[k] <- i-1
  
  while(m[i,c]!=0 && c!=1 && i!=0){
    k <- k+1
    c <- (c-wei[i-1])
    i <- which(m[,c] == m[i-1,c])[1]
    elements[k] <- i-1
  }
  
  value <- round(m[n+1,W+1])
  elements <- sort(elements[which(elements>0)])
  
  result <- list(value=value, elements=elements)
})

# system.time of the dynamic algorithm with different while loop (vectorization)
system.time({
  
  # check input variables
  stopifnot(is.data.frame(x), x>0, length(x)==2, names(x) == c("w","v"))
  stopifnot(is.numeric(W), W>0, length(W)==1)
  
  
  # initialize starting values
  n <- nrow(x)
  val <- x$v
  wei <- x$w

  m <- matrix(nrow = n+1, ncol = W+1)
  m[1,] <- 0
  m[,1] <- 0
  
  for(i in 1:n){    
    for(c in 0:W){ # c = capacity of knapsack
      
      if(wei[i] > c){
        m[i+1,c+1] <- m[i,c+1]
      } else {
        m[i+1,c+1] <- max(m[i,c+1], m[i,c+1 -wei[i]] + val[i])
      }
    }
  } 
  
  c <- c+1  
  i <- which.max(m[,c])
  elements <- length(n)
  k <- 1
  elements[k] <- i-1
  
  while (m[i,c]!=0 && c!=1 && i!=0) {
    k <- k+1
    c <- c-wei[i-1]
    i <- which(m[,c] == m[i-1,c])[1]
    elements[k] <- i-1
}

  
  value <- round(m[n+1,W+1])
  elements <- sort(elements[which(elements>0)])
  
  result <- list(value=value, elements=elements)
})

```
We can notice a little improvement in system.time with the second version of the while loop, in which we used the vectorization method. 

### Greedy algorithm
```{r}

x = knapsack_objects[1:8000,]
W = 3500

profvis({
  
  # initialize starting values
  n <- nrow(x) # number of items
  val <- 0 # current knapsack value
  elements <- rep(0,n)
  ratio <- x$v/x$w      # vector with ratio (value/weight)
  
  i <- 1   
  
  # we find the max value of the weights, we take the position so calculate 
  # the sum and to save the items we are adding
  
  while((sum(x$w[elements]) + x$w[which.max(ratio)]) <= W && any(x$w>0)){
    
    val <- val + x$v[which.max(ratio)] 
    elements[i] <- which.max(ratio)
    ratio[which.max(ratio)] <- 0 # x$w[which.max(ratio)] <- 0
    i <- i + 1
  }
  
  elements <- elements[which(elements>0)]
  result <- list(value=round(val), elements=elements)
})

```
The while-loops is the time-consuming part of this function. 

### System.time Greedy algorithm comparison
```{r}

x = knapsack_objects[1:80000,]
W = 3500

# system.time of greedy algorithm
system.time({
  
  # initialize starting values
  n <- nrow(x) # number of items
  val <- 0 # current knapsack value
  elements <- rep(0,n)
  ratio <- x$v/x$w      # vector with ratio (value/weight)
  
  i <- 1   
  
  # we find the max value of the weights, we take the position so calculate 
  # the sum and to save the items we are adding
  
  while((sum(x$w[elements]) + x$w[which.max(ratio)]) <= W && any(x$w>0)){
    
    val <- val + x$v[which.max(ratio)] 
    elements[i] <- which.max(ratio)
    ratio[which.max(ratio)] <- 0 # x$w[which.max(ratio)] <- 0
    i <- i + 1
  }
  
  elements <- elements[which(elements>0)]
  result <- list(value=round(val), elements=elements)
})

# system.time of greedy algorithm with different while loop
system.time({
  
  # initialize starting values
  n <- nrow(x) # number of items
  val <- 0 # current knapsack value
  elements <- rep(0,n)
  ratio <- x$v/x$w      # vector with ratio (value/weight)
  
  i <- 1   
  
  # we find the max value of the weights, we take the position so calculate 
  # the sum and to save the items we are adding
  
  while(sum(x$w[elements]) + max(x$w[ratio > 0]) <= W && any(x$w > 0)){
    selected_idx <- which.max(ratio)
    val <- val + x$v[selected_idx]
    elements[i] <- selected_idx
    ratio[selected_idx] <- 0
    i <- i + 1
    }
  
  elements <- elements[which(elements>0)]
  result <- list(value=round(val), elements=elements)
})

```
Changed the while loop, memorizing the "which.max()" value instead of calling it three times, gaining a considerable improvement of run time. \newline


Question: What performance gain could you get by trying to improving your code? \newline
Hopefully a faster code. 


## (*) Parallelize brute force search
Question: What performance gain could you get by parallelizing brute force search? \newline
Speeding up the run time of the code. However this is only the case when working with large datasets. In our case the dataset is too small, so the parallelized algorithm takes more time than the normal one. 

